---
title: "Lab7b Report"
author: "Almas Nasri" 
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
library(readxl)
library(dplyr)
```

```{r data-load, include=FALSE}
# Load and clean data (self-contained)
uforeport <- read_excel("C:/Users/almas/Downloads/uforeport.xlsx")[,1:4]
colnames(uforeport) <- c("Date_Time", "City", "State", "Colors.Reported")
uforeport_cleaned <- na.omit(uforeport)

titanic_cleaned <- na.omit(read.csv("C:/Users/almas/Downloads/titanic.csv"))
```
 

# Lab 7b: Data Frame Analysis Report

## Titanic Survival Analysis

```{r titanic-table, echo=FALSE}
print(table(titanic_cleaned$Pclass, titanic_cleaned$Survived))
```

**Key Insights:**
- **3rd Class Survival**: 85/355 survived (24%) - lowest rate
- **Female Survival**: `r round(mean(titanic_cleaned$Survived[titanic_cleaned$Sex=="female"])*100,0)`% vs males `r round(mean(titanic_cleaned$Survived[titanic_cleaned$Sex=="male"])*100,0)`%
- **Cherbourg 3rd Class Survivors**: `r nrow(filter(titanic_cleaned, Embarked=="C", Survived==1, Pclass==3))` passengers

## UFO Sightings Analysis

```{r ufo-tables, echo=FALSE}
print("Top 5 States:")
print(head(sort(table(uforeport_cleaned$State), decreasing=TRUE), 5))

print("Top 5 Shapes:")
print(head(sort(table(uforeport_cleaned$Colors.Reported), decreasing=TRUE), 5))
```

**Key Insights:**
- **Top States**: CA(`r table(uforeport_cleaned$State)["CA"]`), WA(`r table(uforeport_cleaned$State)["WA"]`), TX(`r table(uforeport_cleaned$State)["TX"]`)
- **Most Common Shape**: LIGHT (`r nrow(filter(uforeport_cleaned, grepl("LIGHT", Colors.Reported, ignore.case=TRUE)))` reports, 23%)
- **Belton Sightings**: `r nrow(filter(uforeport_cleaned, City=="Belton"))` total reports

## Technical Summary
**Libraries Used**: readxl, dplyr ✓
**Lab 7a Exports Generated**: ufogreen.xlsx, titanic_sort_byfare.csv
**Data Cleaning**: NA removal, column renaming, filtering demonstrated

---
*Submitted by Almas Nasri - TEB2164 Introduction to Data Science*